\section{Structural modeling} \label{sec:struct}

\subsection{Structural modeling: An overview}

In Sections \ref{sec:causal} and \ref{sec:quasi}, we suggested that researchers minimally consider using diagrams to communicate the basis for their causal inferences, and in Section \ref{sec:mech} we suggested that researchers be more precise in describing how their data permit causal
inferences. This section explores a formal approach to developing a causal model, 
namely the ``structural" approach.
Structural models are empirical models that are derived from theoretical models of behavior.
The term structural model originated with economists and statisticians working at the Cowles Foundation in the 1940s and 1950s.
The earliest structural models used economic models of consumer and producer behavior to derive demand and supply equations.
By adding an equilibrium condition, such as quantities demanded equal quantities supplied, economists obtained a set of mathematical equations that could be used to understand movements in observed prices and quantities. A question then arose as to whether economists could reverse engineer this modeling process and use observed prices and quantities to recover the underlying demand and supply relations.
The models made it clear that the empiricist could only recover estimates of the unobserved demand and supply equations if certain exogenous (instrumental) variables were available. 

The impact of these early models on empirical work in economics encouraged other social scientists to begin using theoretical models to interpret data. 
Structural models have found widest application in situations where causality is an issue, such as the determinants of educational choices, voting, contraception, addiction, and financing decisions. 
Other applications of structural models are discussed in \citet{Reiss:2007ej} and \citet{Reiss:2011go}.

A structural empirical model comprises a theoretical model of the phenomenon of interest and a stochastic model that links the theoretical model to the observed data.
The theoretical model minimally describes who makes decisions, the objectives of decision makers, and constraints on their behavior.
In developing and analyzing the theoretical model, the researcher decides what conditions (variables) matter and what is endogenous and exogenous.
While the theoretical model typically draws on economic principles, it could also be derived from behavioral theories in other fields, such as psychology and sociology.\footnote{Some
researchers refer to any mathematical model fit to data as a structural model. For instance,
one might assume that the number of restatements in an industry follows a Poisson process
and then fit the parameters of the Poisson model using industry-level data on
restatements. We do not view such models as structural because they lack specific behavioral
or institutional components that permit a causal inference. We would classify this approach
as descriptive or statistical modeling.}

Structural models offer a number of benefits for empirical researchers.
First, structural modeling is a process that forces a researcher to make explicit assumptions about what determines behavior and outcomes (i.e., the casual mechanism). 
Second, structural models make it clear what data are needed to identify unobserved parameters and random variables, such as coefficients of risk aversion.
Third, structural models provide a foundation for estimation and inference. 
Finally, structural models facilitate counterfactual analyses, such as what might happen under conditions not observed in the data. 
% I might add a set of limitations here using Welch as a framework -- 
To illustrate these benefits, as well as some of their limitations, we next explore an accounting application.

\subsection{Structural models in accounting: An illustration}
This section develops a model of managerial incentives to misstate accounting information. 
This topic has been the focus of many papers in recent years \citep[see][]{Armstrong:2010jd}.
The key question in this literature is whether certain kinds of managerial incentives increase the tendency for managers to misstate (or attempt to misstate) financial information.
A number of papers hypothesize that tying managers' compensation to the information that they provide will increase their desire to misstate that information.
However, some researchers suggest that, by aligning the long-term interests of shareholders and managers, certain kinds of incentives could actually \emph{reduce} misstatements \citep{Burns:2006ce}.

\citet{Efendi:2007ja} illustrates a fairly typical descriptive empirical paper in this literature. 
\citet[p.\,687]{Efendi:2007ja} estimate a logistic regression with an indicator for restatements as the dependent variable and measures of CEO incentives as independent variables of interest, along with controls for firm size, financial structure, and corporate governance proxies.\footnote{
\citet{Efendi:2007ja} also employ a case-control design, which involves matching firms with restatements with firms without.
We do not focus on that aspect of their research design in our discussion here.}
 
A key assumption implicit in much of this literature is that restatements are a good proxy for actual \emph{mis}statements \citep[e.g.,][]{Efendi:2007ja,Armstrong:2010jd}. 
This assumption is made because in practice accounting researchers only observe misstatements that are detected and corrected by external monitors after the financial statements were issued. 
Examples of these external monitors include whistleblowers, regulators, media, and others \citep[e.g.,][]{Dyck:2010kh}.  For simplicity,  we refer to the actions of these external monitors collectively as ``subsequent investigations."
If subsequent investigations are perfect and detect all misstatements, then there is a one-to-one correspondence between misstatements and restatements.\footnote{There will still be a difference between \emph{attempted} misstatements and actual misstatements due to the external auditor correcting some attempted misstatements.}
Realistically, these subsequent investigations are not perfect, meaning that we need to recognize the difference between misstatements and restatements when estimating the effect of managerial incentives on misstatements.  

In the following analysis, we consider two alternative models of the causal mechanism linking managerial incentives to accounting restatements.
Each model explicitly considers the incentives of the manager and the role of the external auditor. The two models, however, lead to different conclusions about how CEO incentives impact
restatements and misstatements. Not surprisingly, these differences are due to the different
assumptions underlying the models. These differences permit us to illustrate the value of having a theoretical model that can interpret competing empirical estimates.

\subsubsection{Model 1: A non-strategic auditor model}
We assume that firm misstatements are deliberate and are made by a single agent, whom we refer 
to as the `CEO.' 
The CEO is assumed to be rational in the sense that he or she trades off private expected benefits and costs of misstatements when deciding whether to misstate.
Specifically, suppose that the CEO receives a benefit of $B^*$ from the \emph{successful} manipulation of earnings (i.e., a misstatement that is not detected either by the firm's auditors before a report is released or by subsequent investigations). 

Besides the CEO, we assume that  the firm's auditors independently catch and correct attempted misstatements at a constant rate $p_A$ and that the (conditional) probability of subsequent investigations catching a misstatement is $p_I$. Given these assumptions, the probability of a misstatement getting past the firm's  auditor and subsequent investigations is $(1-p_A) \times (1 - p_I)$.
The CEO's expected benefit from a successful misstatement is then
$$ B^* = (1-p_I) \times (1-p_A) \times B$$
where $B$ is a gross benefit to the manager from a misstatement. 

Assume the CEO must exert a fixed cost of effort $C_M$ in order to misstate performance. 
Combining this cost with the manager's expected benefits from of misstatement gives
\begin{equation}\label{bencost}
		y_M^* = 
		  \begin{cases}
				\mbox{Misstate} & \mbox{if }\, (1-p_I) \times (1-p_A) \times B - C_M \ge 0\\
				\mbox{Don't misstate,} & \mbox{otherwise}.
		  \end{cases}.
\end{equation}
This (structural) inequality describes the unobserved misstatement process. 
In general, researchers will not observe the structural objects of
interest: $B$, $C_M$, $p_A$, or $p_I$.

To complete the structural model and recover these objects, 
the researcher must add assumptions that relate the objects to the data available.
Suppose we only observe a (zero-one) indicator variable $y$ for restatements.
These restatements are the result of three decisions:
\begin{enumerate}
\item The manager misstates (or not).
\item The firm auditor detects and corrects an attempted misstatement (or not) .
\item A subsequent investigation detects a misstatement and a restatement occurs (or not).
\end{enumerate}

Mathematically, this sequence can be modeled as
\begin{equation}\label{restateeqn}
 y = I(\mbox{Restate}) = I(y^*_M \ge 0) \, \times\, (1 - I(y^*_A \ge 0)) \, \times\, I(y^*_I \ge 0)
\end{equation}
where $I(\cdot)$ is a zero-one indicator function equaling one when the condition in parentheses is true.
The unobserved variables $y^*_M$, $y^*_A$ and $y^*_I$ reflect the criteria that underlie the CEO's, 
the firm's auditor, and subsequent investigators' decisions.  Notice that equation (\ref{restateeqn}) uses $(1 - I(y^*_A \ge 0))$, an indicator for the firm's auditor missing the misstatement.

Equation (\ref{restateeqn}) somewhat resembles a traditional binary discrete choice model. The easiest
way to see this is to take expectations (from the researcher's standpoint). Assuming the
decision variables are independent,
\begin{equation} \label{equilpr}
\begin{array}{lcl}
 E(y) & = & E\, \left[\; I(y^*_M \ge 0) \, \times\, (1 - I(y^*_A \ge 0)) \, \times\, I(y^*_I \ge 0) \; \right]\\[1em]
 & = &  \mbox{Pr(Misstate)} \times \mbox{Pr(Auditor Misses)} \times
\mbox{Pr(Investigation Finds)}\\[1em]
& = & \beta^* \times (1-p_A) \times p_{I} = \mbox{Pr(Restate)} ,
\end{array}\end{equation}
where $\beta^*$ is the (researcher's) forecasted probability that a misstatement occurs, or
from equation (\ref{bencost})
\begin{equation}\label{betaplus}
\beta^*= \mbox{Pr}\left(\, (1 - p_A)(1 - p_I) B - C_M \ge 0 \,\right)
\end{equation}

At this point, the theory has delivered a structure for relating the \emph{unobserved} probability of a misstatement, $\beta^*$, to the potentially estimable probability of a restatement.
Now we face a familiar structural modeling problem, which is that the model does not anticipate all the reasons why in practice these probabilities might vary across firm accounting statements.
For example, the theory so far does not point to reasons why CEOs might differ in their benefits and costs of misstatements. 
To move theoretical relations closer to the data, researchers typically allow parts of the model to depend on differentiating variables. Often the specifications of these dependencies are ad hoc. 
Empiricists are willing to do this, however, because they believe that it is important to account for practical aspects of the application that the theory does not recognize.

To illustrate this approach, and following suggestions of what might matter from the
accounting research literature, suppose the CEO's unobserved costs and benefits vary as follows:
%
\begin{equation}
\begin{array}{lcl}\label{eqns1}
B & = & b_0 + b_1 \, \mbox{EQUITY} + X_B\beta\\[.5em]
C_M & = & m_0 + m_1 \, \mbox{SALARY} + X_C\gamma + \xi \text{,} % \\[.5em]
\end{array}
\end{equation}
where EQUITY is the fraction of a CEO's total pay that is stock-based compensation, 
the $X_B$ are other observable factors that impact the manager's benefits from misstatements,
SALARY is the CEO's annual base salary, and the $X_C$ are observable factors impacting the CEO's perceived costs of misstatements.\footnote{
For expositional purposes, we assume away $X_B$ and $X_C$ in our analysis.}
The EQUITY variable is intended to capture the idea that, the more a CEO is rewarded for performance, the greater his or her incentive to misstate results so as to increase (perceived) performance.
Thus, we would expect the unknown coefficient $b_1$ to be positive if providing more equity incentives increases the tendency of the CEO to misstate earnings, but expect $b_1 < 0$ if it reduces that tendency.
Similarly, we include the variable SALARY as a driver of the cost of making misstatements, with the idea that if the
CEO is caught misstating, they might lose their job, including salary (and other benefits).
% I want to say that -- The idea is that if the CEO is caught manipulating, the personal penalty or loss borne by the CEO is proportional to his or her compensation. However, this does not seem consistent with the model because the manager always bears this cost regardless of whether he manipulates or not.  Any other justification for this variable come to mind?  Maybe if the salary is high, this means that the CEO is really important and it is more costly for him to fool around with manipulating, as opposed to working on strategy development
Thus, we would expect the unknown coefficient $m_1$ also to be positive.
For now, we leave the other $X$ variables unnamed.

We have no strong theoretical reason for the assumption of linearity. Its motivation is practical, as it facilitates estimation of the model unknowns (as we shall shortly see).\footnote{Another key variable in the above model is the unobserved cost $\xi$.
While it makes sense to say that the researcher cannot measure all misstatement costs, why not also allow for unobserved benefits as well?
The answer here is that adding an unobserved benefit would not really add to the model as it is the net difference that the model is trying to capture.
The sense in which it could matter is if we thought we observed the probabilities $p_A$ and $p_I$.
In this case, we might be able to distinguish between the cost and benefit unobservables based on their variances.}
 
With these assumptions, the probability of a restatement becomes
\begin{equation} \label{restate1}
\mbox{Pr(Restate)} = \theta_0 \mbox{Pr}\left(\, \theta_1 + \theta_2 \mbox{EQUITY} + \theta_3 \mbox{SALARY}  \ge \xi \,\right) .
\end{equation}
The new $\theta$ parameters are functions of the underlying incentive parameters as
follows: $\theta_0=(1-p_A) \times p_I, \theta_1 = (1 - p_A)(1 - p_I) b_0 - m_0, 
\theta_2 = (1 - p_A)(1 - p_I) b_1,$ and $\theta_3 = - m_1$. 
Apart from the scalar multiple $\theta_0$, which can be absorbed into the probability statement (and thus is not identified), this probability model has the form of  a familiar binary choice (e.g., a probit or logit model).
Thus, the value of the structure imposed so far is that it can motivate the application of a familiar statistical model \citep[as in][]{Efendi:2007ja}, as well as explain how the estimated coefficients are potentially connected to quantities that impact the probability of a misstatement.

%Do we want to say something about whether knowledge of the theta parameters can be used to infer the b's.  For examples, the ratio of theta 2 to theta 1 will be b1 divided by b0.  If the model is correct, we might be able to recover the underlying structural parameters.

\subsubsection{Estimation the non-strategic auditor model}

To illustrate how to estimate this structural model, we constructed a data set containing 10,000 firm-year observations on whether or not financial results were restated.
Definitions of the variables in our data set are provided in Table \ref{tab:desc}.
We will discuss the source of the data in more detail later. For those interested,
the data are available from the authors.

The data include variables that have previously been used to model restatements. RESTATE is a zero-one indicator variable for whether a firm restated (RESTATE = 1) their financial results in a given year.  
%TODO: Provide a reference for BIG4 in this kind of regression.
The variable BIG4 also is a zero-one indicator for whether the firm's auditor is one of the four largest U.S. accounting firms. It is included in the specifications because Big 4 auditing firms might have more accounting expertise and this expertise might make them more likely to catch misstatements. 
Similarly, the corporate governance literature suggests that board oversight from directors with accounting or finance backgrounds reduces the likelihood of misstatements. We proxy this possibility with FINDIREC, the percentage of directors that have professional accounting or finance backgrounds.
%TODO: Get refs.
Finally, the variables INT and SEG are included to capture the complexity and costs of audits. 
%TODO: From AAZ: Link to variable definitions.
%TODO: Get reference for audit model.
Specifically, INT is a zero-one indicator for whether the firm does a majority of its business outside the U.S.. We believe that international companies have higher auditing costs. Similarly SEG is a count of the firm's two-digit Standard Industrial Classification business segments. We believe that more segments likely will increase the costs of auditing. 

Table \ref{tab:desc} reports descriptive statistics for our sample. 
CEOs on average receive about one million dollars in base pay and their equity-related pay averages 45\% of their total pay.
About three-quarters of the sample has a Big 4 accounting firm as its auditor. 
The fraction of directors with financial expertise is less than ten percent. 
The average firm has about 1.5 business segments and is primarily based in the United States.

Table \ref{tab:logit} reports the results of logit regressions in which the dependent variable is the restatement indicator variable. These specifications parallel prior descriptive statistical models that correlate restatements with other variables that might impact misstatements.
The table contains both a simple specification containing an intercept along with the two CEO 
compensation variables, 
and a more intricate specification involving the other variables in the dataset.
For each specification, we report the estimated coefficients of the logit and the corresponding marginal effects evaluated at the sample means of the exogenous variables.

The results for the pay coefficients in both specifications run counter to those the previous accounting literature might predict and counter to those predicted by the structural model 
that assumes the benefit coefficient on equity pay, $b_1$, is greater than zero.
Specifically, more base pay is associated with more restatements, while more equity-based 
compensation is associated with fewer restatements.

Besides the intercepts and the EQUITY and SALARY coefficients, the only other coefficients that are statistically significant are those on INT and SEG.
While we can say (descriptively) that INT and SEG are associated with higher restatement rates, unless we take a position on how they enter $X_C$ or $X_B$, it is difficult to interpret whether these signs make sense.

The question we now address is what to make of the fact that the coefficients on EQUITY seem inconsistent with our informal arguments and with the prediction from our structural model 
that assumes $b_1>0$. One possible interpretation of this finding is that the our prior
beliefs about incentives were wrong. Another possibility is that the measures we employ and
the functional forms assumed are incorrect, which then leads to spurious results. Yet another
possibility is that our theory of misstatements is incorrect. It is this later possibility
we consider here.

% PCR: Do we need this paragraph? Did we really consult colleagues? (The responses suggest we may have done so.) -IDG
% We consulted colleagues in the profession for their opinions.
% One set attributed the outcome to data errors or sampling issues. 
% Another set said the logit model was obviously flawed because it did not include all relevant accounting variables. (High on the list were measures of recent accrual activity, use of off balance sheet transactions and firm performance variables.) 
% Their thinking was, had we included these, the signs on the pay coefficients might be different. 
% Mentioned less often was that the idea that an important difference between restatements and misstatements is the role of subsequent investigations. 
% There is nothing in the model or variable list that would help predict the intensiveness of outside scrutiny (or indeed of scrutiny by the firm's external auditors).

\subsubsection{Model 2: A strategic auditor model}
A key weakness of the previous model is that it ignores the incentives of the  external auditor.
According to PCAOB guidance in Auditing Standard No. 12, assessment of the risk of material misstatement should take into account ``incentive compensation arrangements."
Similarly, Auditing Standard No. 8 suggests that audit effort should increase if risk is higher.
To make the model richer in a manner consistent with these institutional details, we assume that auditors trade off the costs of audit effort against the reputational losses they might incur should they miss a managerial misstatement that is subsequently detected.\footnote{
Here we have in mind the findings of \citet{Dyck:2010kh} who show that many egregious forms of misstatements are detected subsequently by employees, directors, regulators, and the media.} 

In the previous model, the firm's auditor impacted the manager's misstatement benefits through $p_A$ (which is assumed to be constant). 
Suppose that $p_A$ is in fact a choice variable for the firm's auditor. 
To make matters simple, suppose that the auditor detects manipulation with probability $p_{AH}$ if they exert high effort and  otherwise they detect manipulation with the lower probability $p_{AL}$. 
Let the cost of high effort be a fixed cost $C_A > 0$. 
Without loss of generality suppose the cost of low effort is zero. 
When deciding whether to audit with high or low effort, the auditor perceives a cost to its reputation, $C_R$, due to not detecting a misstatement that is caught by subsequent investigations. 
This structure implies that the total cost of high effort to the auditor is $C_A + (1-p_{AH}) \times p_I \times C_R$ or the cost of high effort plus the expected cost of missing a misstatement that is subsequently caught with probability $p_I$. 
The total expected cost of
low effort is similarly equal to $(1-p_{AL}) \times p_I \times C_R$. 

To complete this new model, we need to make an (equilibrium) assumption about how the CEO and firm auditor interact. 
Following the literature, we assume that the two simultaneously and independently make decisions, and that their strategies form a Nash equilibrium.
That is, we assume the players' strategies are such that they optimize their objectives  taking the actions of the other players as fixed. This means that in a Nash equilibrium, the players are taking actions that they cannot unilaterally improve upon.

In this type of auditing game, the Nash equilibrium has the CEO and the auditor playing mixed (randomized) strategies.
That is, the auditor will independently exert high effort with probability $\alpha^*$ and the CEO independently misstates with probability $\beta^*$. 
These probabilities are such that each party has no incentive to change strategy. 
That is:
%\vspace{-3mm}
\begin{enumerate}
\item the CEO
is indifferent between misstating and not misstating, or:
\begin{equation}\label{manager}
(1 - p_A^*)(1 - p_I) B - C_M = 0 
\end{equation}
where $p_A^* = \alpha^* p_{AH }+ (1-\alpha^*) p_{AL}$ is the equilibrium 
probability a misstatement is detected; and,
\item the auditor is indifferent between exerting high and low effort, or
$$ \beta^* (1-p_{AH}) p_I C_R + C_A = \beta^* (1-p_{AL}) p_I C_R .$$
\end{enumerate}

Solving these two equations for the equilibrium probabilities $\alpha^*$ and $\beta^*$
yields:
\begin{equation}\label{equilstrat}
\begin{array}{lcl}
  \alpha^* &= & \dfrac{ ( 1 - p_{AL}) (1 - p_I) B- C_M}{ (1 - p_I) (p_{AH}-p_{AL}) B}\\[1.5em]
  \beta^* &= & \dfrac{C_A}{(p_{AH}-p_{AL}) p_I C_R}  
\end{array}
\end{equation}
From these equations, we can calculate the equilibrium probability of a restatement\footnote{
As part of the solution, we require $\alpha^*$ and $\beta^*$ to be probabilities between
zero and one. 
This is true provided $C_R$ and $B$ satisfy the inequalities
$ C_R > \frac{C_A}{(p_{AH}-p_{AL})p_I} $
and 
$ B > \frac{C_M}{1 - p_I}  $.}
\begin{equation} \label{equilpr1}
\begin{array}{lcl}
\mbox{Pr(Restate)} & = &  \mbox{Pr(Misstate)} \times \mbox{Pr(Auditor Misses)} \times
\mbox{Pr(Investigation Finds)}\\[1em]
& = & \beta^* \times (1-p_A^*) \times p_I
\end{array}\end{equation}
This equation tells us how the probability of a restatement is related to
the unobserved frequency of misstatements. In particular, if we knew the frequency with which auditors and subsequent investigations caught misstatements, we could easily link the two. Otherwise,
we would have to estimate these probabilities (or make assumptions about them).

Substituting the equilibrium strategies (\ref{equilstrat}) into (\ref{equilpr1}) yields
\begin{equation} \label{equilpr2}
\begin{array}{lcl}
\mbox{Pr(Restate)}& = &  \dfrac{C_AC_M(1-p_{AL})}{(p_{AH}-p_{AL})(1-p_I)C_RB}.
\end{array}\end{equation}
We now are in a position to use the theory to help interpret the conflicting logistic regression results
in Table 3. 

Equation (\ref{equilpr2}) shows that the presence of a strategic external auditor
changes how the CEO's incentives impact the probability of a restatement.\footnote{Notice that
the probability statement in equation  (\ref{equilpr2})  differs from that in equation (\ref{restate1}).
The probability statement in equation  (\ref{equilpr2}) reflects the randomness of the strategies, whereas in equation (\ref{restate1}) it reflects variables the researcher does not observe.} 
Partial derivatives of equation (\ref{equilpr2}) show that the restatement probability is:
\begin{itemize}
\item Decreasing in the benefit $B$ that the CEO enjoys from misstatement.
\item Increasing in the personal cost of manipulation $C_M$ incurred by the CEO .
\item Decreasing in the reputational cost $C_R$ incurred by the external auditor.
\item Increasing in the cost of high effort $C_A$ incurred by the external auditor.
\end{itemize}

Thus, in contrast to the model with a non-strategic auditor, increasing the benefit that managers enjoy from misstatement, or decreasing the misstatement cost, leads to fewer restatements being observed by researchers.
These two effects might explain the negative sign on EQUITY and the positive sign on SALARY observed in the previous logit results. 
% Thus, this structural model has the potential to rationalize patterns observed in the data with beli
%NOTE: The logit results and an assumption that b_1 < 0 *also* rationalizes the "patterns observed in the data."

To have a better sense of how one might connect the strategic auditor theory to the
logistic models in Table \ref{tab:logit}, suppose, similar to the motivation for equation (\ref{eqns1}), that 
\begin{equation}\begin{array}{lcl}\label{eqns2}
B & = & b_0 + b_1 \, \mbox{EQUITY} \\[.5em]
C_M & = & m_0 + m_1 \, \mbox{SALARY} \\[.5em]
C_A & = & a_0 + a_1 \, \mbox{INT} + a_2 \, \mbox{SEG}\\[.5em]
C_R & = & r_0, \quad p_{AH}   =  p_0, \; \mbox{ and } \; p_{AL}  =  v_0 \\[.5em]
\end{array}
\end{equation}
where $ a_0, a_1, a_2, r_0, b_0, p_0$ and $v_0$ are constant parameters. 
Inserting these expressions into the expected restatement rate (\ref{equilpr2}) gives
\begin{align*} \label{equilpr3}
\mbox{Pr(Restate)} & =   \dfrac{C_AC_M(1-p_{AL})}{(p_{AH}-p_{AL})(1-p_I)C_RB} \\
&= \dfrac{(1-v_0)(a_0 + a_1 \, \mbox{INT} + a_2 \, \mbox{SEG})(m_0 + m_1 \, \mbox{SALARY})}
{(p_0-v_0)r_0(b_0 + b_1 \, \mbox{EQUITY})}
\end{align*}
\begin{equation}\label{equilpr4}
 =  \dfrac{\theta_0 + \theta_1\mbox{\small INT} + \theta_2\mbox{\small SEG} + \theta_3\mbox{\small SALARY}
+ \theta_4\mbox{\small INT} \times \mbox{\small SALARY}+ \theta_5\mbox{\small SEG} \times \mbox{\small SALARY}}
{1 +  \theta_6\mbox{\small EQUITY}}
\end{equation}
Notice that the $\theta$'s absorb unknown quantities such as $r_0$ and $p_0$, and that the denominator intercept
is normalized to one. This last restriction is required to identify the ratio of the two linear functions.

Although this model does not have a logit form, it is potentially estimable using a nonlinear
estimation method such as generalized method of moments (GMM).\footnote{Other estimation 
strategies are possible, but we do not consider them here.}
GMM attempts to match sample moments to what the structural model implies 
these moments should be. 
For example, an obvious sample moment would be the average restatement rate in the sample.
The corresponding theoretical moment would be the probability expression in equation (\ref{equilpr4}).
In our estimations,  we use sample moments of the form:
\begin{equation}\label{gmoment}
\mathcal{M}_j = \sum_{i=1}^{10,000} \; X_{ji}^\prime \left[ \; \mbox{RESTATE}_{i} - \mbox{Pr(Restate}_i\mbox{)} \; \right]. 
\end{equation}
where Pr(Restate) comes from equation (\ref{equilpr4}).\footnote{
To ensure that the model parameters imply restatement probabilities between zero and one, we add a penalty function to the GMM objective function.
This penalty increases with the number of estimated probabilities below zero or above one.
For most replications this penalty is immaterial to the results obtained.}
Practically, we need at least as many moments as we have $\theta$ parameters to estimate (there are seven $\theta$'s in the model). The $X_j$ used in the moments include all explanatory variables,
plus some interactions. (See Table 3 for a list.) Again, to illustrate how we estimate
the $\theta$ parameters of equation (\ref{equilpr4}), one of the $X$'s is
a dummy variable for whether the firm is an international company. The corresponding moment equation 
in (\ref{gmoment}) seeks to match international companies' average restatement rate to the model's prediction for that rate.

Table \ref{tab:gmm} reports the results of estimating the new (strategic auditor) structural model. 
The results show that in this particular case, even without sample information on the unobserved probabilities $p_A$ and $p_I$, we can recover estimates of the model parameters up to normalizations.
For instance, the coefficient ratio $\theta_4/\theta_1$ estimates the ratio of cost parameters $m_1/m_0$.
The parameter $m_1$ is the cost coefficient on SALARY and 
$m_0 > 0$ is a fixed cost of manipulation. The sign of $\theta_4/\theta_1$ thus reveals the sign of $m_1$.
From the theory, we expect the sign to be positive, and this is what we find in the estimation results.\footnote{The same is not true of $\theta_3/\theta_0$.}

Similarly, $\theta_6$ equals the (scaled) misstatement benefit coefficient on the EQUITY variable.
Recall that the descriptive logit regression coefficients in Table 2  suggest EQUITY has a negative affect on misstatements.  In contrast, we now find the expected positive relation because we explicitly model the difference between misstatements and restatements in our structural estimation. 

Table 3 contains three different sets of estimates. Column
(1) are for an exactly identified model where there are
as many instruments as parameters to estimate. Column (2)
presents estimates for a over-identified model in which
there are more moments than parameters to estimate. Column
(3) presents constrained GMM estimates, where the constraints
are motivated by the fact that pairs of ratios of the
$\theta$ coefficients are equal (e.g., $\theta_3/\theta_0
= \theta_4/\theta_1$). The three sets of estimates are
similar, with the constrained estimates seemingly yielding
much more statistical precision.


\subsubsection{Implications of structural modeling analyses}
From the discussion above, it seems that there are (at least) two alternative explanations (or hypotheses) for the results we find.
One hypothesis is that the process generating the data is best modeled with a non-strategic auditor and that the effect of EQUITY on incentives to misstate is either negative (or perhaps zero).
The support for this hypothesis comes from Table \ref{tab:logit}, which is an appropriate regression analysis for the model with a non-strategic auditor, where a negative (and weakly statistically significant) coefficient on EQUITY is found.
However a second, and in our view a considerably more plausible, hypothesis is that the process generating the data is best modeled with a strategic auditor and that the effect of EQUITY on incentives to misstate is positive (or perhaps zero).
The support for this hypothesis comes from Table \ref{tab:gmm}, which is predicated on the model with a strategic auditor, and where a positive coefficient on $b_1$ (the parameter linking EQUITY to benefits from misstatement) is found.

The point of this discussion is not to resolve the debate regarding the effect of incentives on misstatements. 
Rather, the goal is to illustrate the necessity of having an underlying structural model of the process by which the data we observe were generated.
The importance of such models was illustrated in Sections \ref{sec:causal} and \ref{sec:mech}, where we used causal diagrams as a kind of (non-parametric) causal model.
Here we have shown more can be inferred from a formal model tied to behavioral assumptions.

Not only does a structural model enable us to derive sharper predictions regarding the relations between variables for various parameterizations, but it also provides a basis for actually estimating those relations. 
In particular, the comparative statistics of the model shed light on the difference between restatements and misstatements, and what assumptions (e.g., a strategic or a non-strategic auditor) and data were needed to draw inferences about misstatements from restatements. 
Additionally, we were able to recover some of the primitive parameters impacting incentives for managers to misstate results, as well as perform counterfactual analyses. Finally, although structural modeling does not allow us to completely resolve questions of causality, if the model is based on reasonable assumptions and has a close fit to the data, we arguably have better insight into the likely causal relations underlying the phenomenon being examined. 
% While there is the potential for disappointment in the simplistic theory we used, we see room for improving models as an opportunity rather than a defining limitation.

\subsubsection{Counterfactuals}

While the coefficient magnitudes only allow us to infer relative
coefficient magnitudes and signs, we nevertheless can use the model to perform counterfactual calculations. A counterfactual calculation
asks what is the consequence of changing some parameters or variables
while holding others fixed.
There are many different counterfactuals that could be considered. 
For illustrative purposes, we evaluate what would happen to misstatements and restatements if we do away with equity-based compensation and nothing else changes in the model.
The value of having an equilibrium model to analyze this change is that we explicitly allow the auditing process to adjust to the removal of CEO incentives to misstate. 
From the equilibrium strategies in equation (\ref{equilstrat}), we see that removing equity-based pay ($b_1=0$ or EQUITY=0) does not change the equilibrium frequency of misstatements (i.e., $\beta^*$), but does change the frequency of high effort auditing ($\alpha^*$).
Inserting the estimated parameters into Equation (\ref{equilpr3}), we find
$$ \dfrac{\mbox{Pr(Restate }\vert \mbox{ No Equity)}}{\mbox{Pr(Restate }\vert \mbox{ Equity)}}=\dfrac{\beta^* \times (1-p_A^{**}) \times p_{I}}
{\beta^* \times (1-p_A^{*}) \times p_{I}} = \dfrac{(1-p_A^{**})}{(1-p_A^*)} = 1.24.$$
This result tells us that the restatement rate would increase by 24\% (or from 9.99\% to 12.36\%) if equity-based incentives were removed or
did not impact the benefits of misstatements. 
The fact that the restatement rate goes up may at first seem somewhat odd given that the benefits to the CEOs have fallen. 
The model, however, shows that the increase  comes about because the auditors exert \emph{less} effort in detecting misstatements, thereby catching fewer, leaving more for subsequent investigations to detect.

\subsection{Limitations of structural models}

When we discuss structural models with colleagues, we
sometimes encounter two negative reactions. One objection
appears to be that these models are both too complicated
and unrealistic to yield much insight on practical accounting
research questions. The other is related to a preconception
that structural modelers believe that all empirical work needs to be structural.

To be clear, we do not believe that all accounting researchers need estimate structural models. Indeed, in many situations, descriptive models can be more informative. Further, no structural modeling exercise should go forward unless the researcher is convinced that the benefits of a structural model would outweigh the substantial costs entailed in developing and estimating a structural model. 

We have mentioned some of the benefits of structural models,
particularly when it comes to having a clear basis for making
causal statements. What are the costs to developing and estimating structural models? 
First, structural models can be technically demanding to develop. 
Additionally, when constructing a theoretical model that can be taken the data, the empirical researcher is typically forced to make simplifications that a pure theorist might never make and that other empiricists criticize as unrealistic. Further, as we saw in the restatement model, the structural modeler often is in the uncomfortable position of ex post adding features, such as covariates, to the model that make it more in line with the
realities of the application.

While these costs are important to recognize, it is also important
to realize that without structural models, there likely will
continue to be a substantial divide between theoretical and empirical research in accounting.
With few exceptions, theoretical accounting researchers do not explain how to map the specifics of their models to data. 
In many cases, extant theory is not sufficient to motivate the hypotheses tested by empirical researchers.
%For example, \citet{Huang:2014cs} study the effect ``tone management'' on capital market outcomes. Developing a formal theory of the relation between firm performance, managerial psychological states, and measures of tone would be a complex undertaking involving economics, psychology, and linguistics.Building on such a (hypothetical) foundation to solve the complex game involving managers and capital markets would be extremely ambitious.
Consistent with the existence of this gap, few empirical research papers in accounting rely on formal theoretical models to motivate their hypotheses. 
Often when empirical researchers \emph{do} rely on theoretical papers to motivate hypotheses, the predictions claimed to be derived from those papers have little obvious connection with the actual content of those papers.
% For example, \citet{Hollander:2010jg} suggests that ``one camp [of analytical research] argues that firms with good information quality will issue less expansive disclosures because information asymmetry is lower in such firms \citep[e.g.][]{Verrecchia:1983}," notwithstanding specific results in \citet{Verrecchia:1983} that support this claim.
% Another camp shows that as information quality increases, which, in turn, increases the quality of the manager's information, managers are incentivized to disclose more because investors will deem such disclosures more credible (Verrecchia [1990b])
Instead, almost all empirical research papers in accounting use more informal, verbal approaches to hypothesis development.

Second, although structural models can in principle make it clear what a researcher is assuming about causality, it is incumbent on the structural modeler to make the model's causal relations clear. One way to do this is to provide a causal diagram. To illustrate, Figure \ref{fig:audit} provides a causal diagram representing the strategic auditor restatement model.\footnote{
While the mixed-strategy of our model has $\beta$ not being a function of $B$ or $C_M$ and $\alpha$ not being a function of $C_A$ or $C_R$, we have retained these links as being plausible in a more general model.}
As can be seen, we are assuming that $p_I$ is independent of EQUITY. 
But is quite plausible that these investigations are conducted (in part) by a regulator who is as strategic as the auditor in our model, thus giving rise to a link between EQUITY and $p_I$.
We also assume that EQUITY is exogenous, whereas it is plausibly related to the complexity of the business, which may also affect the cost of auditing.
These links could be added to the structural model, albeit at some cost.
Evaluation of their accuracy then could be made via in- or out-of-sample goodness-of-fit tests.

Third, just because a researcher can write down a theoretical model and estimate it does not make the empirical model ``right."
Clearly there is a risk of incorrect causal inferences being drawn from estimation of a structural model based on faulty assumptions.
In our analyses, we in fact generated the data ourselves. \footnote{The parameter values used to generate the data are: $a_0 = 0.5 , a_1 = 3.5 , a_2 = 3.5,
m_0 = 7, m_1 = 1.5, b_0 = 20, b_1 = 10, p_0 =0.75,
v_0 = 0.05, p_I = 0.45$ and $r_0 = 60.$}That is, the data are not from actual firms, but instead were generate by computer using the strategic auditor model. We chose to do this
so that we could more convincingly show that structural models
are capable of recovering behavioral information provided the
model is correct. This can be seen in Table 3 by comparing the
"True" coefficients in the last column that were used to generate the data to the other columns. Because we also used the same
data to estimate the logits in Table 2, we are able to establish
that if we used the wrong (non-strategic) auditor structural model, we would have been led astray in our inferences about the effect of EQUITY.

In practice we do not have this kind of insight into the correct process that generated the data. Hence, the researcher will need to weigh models based upon how well the model's assumptions match the practical and institutional realities of the phenomena being studied.
%TODO: Tone down or eliminate?
% \cite{Nikolaev:2014er} provide an analysis of misreporting and accounting quality.
% The critical assumption in \citet[p.\,6]{Nikolaev:2014er} is the ``idea that both operating cash flow and earnings can be viewed as noisy measures of underlying economic performance."
%DL: Not sure that I understand the discussion here on Nikolaev
%But the fact that cash flow from operations is intended to measure something quite different (e.g., investing cash flows affect subsequent earnings quite differently from operating cash flows) suggests that this may not be the best assumption with which to identify a structural model.
Despite these challenges, we believe that there is significant value in making the theory underlying empirical research transparent and rigorous.\footnote{The use of structural models in accounting research has been fairly limited to date.  
Recent examples include \citet{Gerakos:2013cl}, \citet{Zakolyukina:2015aa}, and \citet{Bertomeu:2015aa}.  
%NOTE: I don't think this applies to Nikolaev, so I took that paper out.
These three papers model an institutionally rich problem, estimate the derived model, provide estimates for important structural parameters, and also give interesting counterfactuals based on their theoretical models. 
We view these papers as useful initial steps in applying structural approaches to accounting research questions.}

%\subsection{Structural models in accounting research} The use of structural models in accounting research has been fairly limited to date. There are certainly examples of regression models being derived (or perhaps influenced) by theoretical models. For example, \citet{Lambert:1987} use the model of \citet{Holmstrom:1979aa} (and a variety of simplifying assumptions) to help specify a regression function linking CEO compensation to firm performance. While somewhat structural in orientation, the approach suppresses the fundamental causal mechanism associated with compensation decisions and does little to actually estimate structural parameters of the model of \citet{Holmstrom:1979aa}.

% I think Gerakos's paper has a manager who seeks to smooth earnings.\citet{Gerakos:2013cl} assume a simple stochastic process for earnings and assume that managers seek to smooth earnings \citet[p.\,57]{Gerakos:2013cl} find evidence that ``unmanipulated earnings are more correlated with contemporaneous returns and have higher volatility than reported earnings."Similar to our simple model above, \citet{Zakolyukina:2015aa} is concerned with how equity incentives motivate managers to manipulate earnings.  \citet{Bertomeu:2015aa} examine management forecasts using a formal disclosure model to estimate whether managers strategically withhold information from shareholders.These three papers model an institutionally rich problem, estimate the derived model, provide estimates for important structural parameters, and also give interesting counterfactuals based on their theoretical models. We view these papers as useful initial steps in applying structural approaches to accounting research questions.
% \footnote{Maybe review some structural finance papers -- Luke Taylor on CEO labor markets, Whited on debt?, Terry jmp about "meet or beat" and macroeconomic research and development.} % Is this footnote for the editors or for us?
