\documentclass[14pt]{article}
\usepackage[marginratio=1:1]{geometry}  % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper} % ... or a4paper or a5paper or ... 
%\geometry{landscape}  % Activate for for rotated page geometry
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
%\usepackage{amsfonts}
\usepackage{soul}
\usepackage{palatino}
\usepackage{natbib}
\usepackage{hyperref} 
\usepackage{amsmath} %Never write a paper without using amsmath for its many new commands 
\usepackage{amssymb} %Some extra symbols 
\usepackage{makeidx} %If you want to generate an index, automatically 
\usepackage{graphicx} %If you want to include postscript graphics 
\setlength{\parindent}{0em}		% No parag indentation
\setlength{\parskip}{1.5ex}	
\newcommand\T{\rule[0em]{0pt}{1.5em}} % Top strut
\newcommand\B{\rule[-1em]{0pt}{0pt}} % Bottom strut
% \usepackage{paralist}

\begin{document}
\large
 
\include{structural}

\end{document}


Empiricists in economics, marketing, finance and other business-related disciplines 
use structural empirical models to draw causal inferences, measure unobserved theoretical
objects and perform counterfactual ``what if" analyses.

changes how the probability of restatement varies with .

From our previous
discussion of CEO incentives, we expect $\theta_2 > 0$ and $\theta_3 < 0$.

\vfill\eject

Equation (\ref{equilpr}) reveals some interesting insights about how the {\it observed}
probability of restatement is related to the {\it unobserved} probability of misstatement.
Specifically
$$ \mbox{Pr(Restate)} \ne  \mbox{Pr(Misstate)} $$
unless $ \mbox{Pr(Internal Auditor Misses)} = \mbox{Pr(External Auditor Finds)} = 1.$ That is,  the unobserved rate of misstatements is the observed restatement rate
if the internal auditor finds no misstatements and the external auditor finds all misstatements. Since this is highly unlikely, to estimate the probability of
misstatements, we must recover the unobserved probabilities that the
internal and external auditors will catch misstatements.




The variables $B, C_M, C_R$ and $C_A$ are all unobserved, yet they are objects of
interest. Unless they vary across firms and/or time, there is no way to recover them
from the data on restatements. This is where structural modelers must add more
assumptions in order to estimate quantities of interest. We do, however, have the
prior literature that can suggest how these might vary. Our previous discussion suggested
CEOBONUS increases manager benefits to misstatement,  $B$. CEOPAY increases the
unobserved costs to the manager of misstatements, $C_M$. Further, we hypothesized
that cost of high effort for the auditor depends on whether the auditor is a Big 4 firm, with
Big 4 firms having lower costs.

A simple specification of how these variables affect the unobservables is
\begin{equation}\begin{array}{lcl}\label{eqns1}
B & = & b_0 + b_1 \, \mbox{CEOBONUS}\\[.5em]
C_m & = & m_0 + m_1 \, \mbox{CEOPAY}\\[.5em]
C_A & = & a_0 + a_1 \, \mbox{INTERNAT} + a_2 \, \mbox{SEGMENTS}\\[.5em]
C_R & = & r_0
\end{array}
\end{equation}
where the constants $b_0, b_1, m_0, m_1, a_0, a_1, a_2$ and $r_0$ are unknown parameters to
be estimated. The last equality reflects the lack of observable covariates that might
cause the auditor reputation cost to vary across observations.

The equations for misstatement and auditing probabilities also depend on the
probabilities $p_I, p_{IH}$ and $p_{IL}$. The presence of finance directors increases
the chance that misstatements will be found, or
$$ p_I = \rho_0 + \rho_1 \, \mbox{FINDIRECT}$$
and the presence of a Big 4 auditor might result in an increase in the high-effort
audit probability
$$ p_{IH} = \tau_0 + \tau_1 \, \mbox{BIG4}.$$

At this point, the observable exogenous variables all enter the theoretical model.
Aside from introducing unobservables, the model is now complete and potentially
estimable. Before turning to estimation issues, 
a natural question at this point is how did we know that these functions
would be linear and only include these variables. The answer is we do not know. This is
a part of structural models that can be criticized as not directly guided by theory. On the
other hand theory rarely recognizes all the factors that might come into play in real data.
If the empiricist is to identify these quantities, some covariation with observables
across observations is necessary.

Another thing to notice before turning to estimation issues is that these specifications
are capable of explaining what we see in the Sample 1 and Sample 2 regressions.
Fixing the internal auditor's behavior, the equation that predicts manager misstatements\\
\begin{equation}
\mbox{Misstate if:}  \qquad (1 - p_{I})(1 - p_G) B - C_M > 0 
\end{equation}
suggests expected misstatements are\\
$$ \mbox{Pr(Misstate)} =  \mbox{Pr}((1 - p_{I})(1 - p_G) B > C_M)$$
Assuming for the moment that the auditing probabilities are constants\\
\begin{equation}\label{misstate}
 \mbox{Pr(Restate)} = (1 - p_{I})\, p_G\mbox{Pr(Misstate)}.
\end{equation}
Because $B$ is increasing in manager incentive pay (CEOBONUS), the model predicts
it will increase restatements, and in Sample 1 indeed it does. Similarly manager base pay (CEOPAY)
should decrease restatements and misstatements, and indeed they do in Sample 1.
On the other hand, when $p_A$ is not a constant, but depends on the auditor's
anticipation of the manager's costs and benefits to misstate, we find in equilibrium 
(\ref{equilpr}) that the opposite is true -- we expect a {\it negative} association on 
average between CEOBONUS and misstatements, and a {\it positive} association with 
CEOPAY. This is what is observed in the Sample 2 regressions. Thus the structural
model can be fit to either data set and potentially rationalize the differences between
the two seen in the regression results.