---
title: "RDD with an endogenous discontinuity"
author: "Ian Gow"
date: "December 8, 2014"
bibliography: papers.bib
output:
  pdf_document:
    highlight: pygments
header-includes:
- \usepackage{bbm}
- \usepackage{palatino}
---

Regression discontinuity designs (RDD) have emerged as an approach for warranted causal inference in several settings in accounting research. In this note, we show that if the discontinuity is endogenous (i.e., determined with reference to the outcome variable of interest), the local nature of RDD estimates may severely compromise their usefulness for understanding the treatment effect for a broader class of firms.

To understand this issue,we examine an idealized version of the setting studied by @Iliev:2010, who uses the market capitalization cutoff of \$75 million in studying the effect of SOX Section 404.\footnote{Note that Iliev (2010) is careful not to include firms with market capitalization far from \$75 million and emphasizes that his estimates apply only to ``small firms.''} 
Consistent with this limitation, our analysis suggests that RDD can provide little insight into the broader question of whether SOX is value-creating for all firms subject to it.

Suppose that a regular observes that the (gross) benefit of a treatment $t$ for firm $i$ equals $B x_i + \epsilon_i$, where $B$ is a constant, $x_i$ is some firm-level characteristic (such as sales or market capitalization) and $\epsilon_i$ is a noise term uncorrelated with $x_i$.
The cost of adopting this treatment is $C$.


Note that the net benefits of the treatment are positive if $B x_i + \epsilon_i - C > 0$ and the expected benefits are positive if $B x_i  > C$.
Thus, if the regulation requires treatment only for firms with $x_i > C/B$, the expected net benefit for these firms will be positive and, for the firms with $x_i < C/B$, the net benefit of treatment would be negative.\footnote{We put aside for now the explanation for why firms do not choose to implement this measure independent of the regulation requiring it.}
Suppose we wanted to run an event study on the date that the regulation was announced.
Suppose that the return on the date of implementation \emph{outside of the effect of the regulation} has expected value of zero and standard deviation of $\sigma_{\eta}$.
So announcement-date return for firm $i$ equals 
$\mathbb{I}_{x_i>C/B} \left( B x_i + \epsilon_i - C \right) + \eta_i$,
where $\mathbb{I}_{x_i>C/B}$ indicates that the regulation applies to firm $i$.

## Simulation 

```{r parameters, include=FALSE}
B <- 0.002
C <- 0.1
N <- 1000
x_max <- 100
sigma_eta <- 0.025
sigma_epsilon <- 0.01 
```

To examine the efficacy of RDD, we simulate `r N` observations using values $B = `r B`$, $C = `r C`$, $\sigma_{\eta} = `r sigma_eta`$, $\sigma_{\epsilon} = `r sigma_epsilon`$, and $x \sim U[0, `r x_max`]$.


```{r simulation}
x <- runif(n = N, 0, x_max)
epsilon <- rnorm(n = N, sd=sigma_epsilon)
eta <-  rnorm(n = N, mean = 0, sd = sigma_eta)
cutoff <- C/B
reaction <- (x > cutoff) * (B*x + epsilon - C) + eta
```

```{r estimation, include=FALSE}
    source("http://iangow.me/code/rd_opt_bw.R")

    ### (4) Calculate the treatment effect and standard errors.
    y <- reaction
    c <- cutoff

    # First, get bandwidth
    h <- rd_opt_bw(y, x, c)
    
    ## Weights based on triangular kernel
    wgts <- (1 - abs(x-c) / h) * (abs(x-c) <= h)
    local.lm <- lm(y ~ (x >=c) * I(x-c) , weights = wgts)    

    rd.est <- coef(local.lm)[[2]]
    sd.est <- sqrt(vcov(local.lm)[2, 2])
     
    ## Output
    out <- c(h, rd.est, sd.est, rd.est/sd.est)
    names(out) <- c("Optimal Bandwidth", "RD Estimate", 
                    "Standard Error", "t-statistic")
    print(out)
```
We estimate the treatement effect using local linear regression, the triangular kernel 
and the [Imbens-Kalyanaraman bandwidth](http://iangow.me/code/rd_opt_bw.R).
The estimated treatment effect is $`r formatC(out[2], digits=3, format='f')`$
($t$-statistic of $`r formatC(out[4], digits=3, format='f')`$).
See below a plot of the data with fitted values.\footnote{Estimates are based on locally linear kernel estimates for each value of $X$ using the bandwidth calculated above. While the bandwidth is optimized at the cutoff, there seems to be merit in visually inspecting the resulting kernel estimates for all values of $X$.}
But note that the average treatment effect for the treated equals the average value of 
$B x_i + \epsilon_i - C$ for the treated firms, which equals 
$`r formatC(sum((x > cutoff) * (B*x + epsilon - C))/sum(x>cutoff), format='f', digits=3)`$
(close to the theoretical expected value of $`r (x_max+C/B)/2*B - C`$).

```{r plot, warning=FALSE, echo=FALSE, fig.width=8, fig.pos='h'}
    # A function to estimate local linear regression around a point (x_i) using
    # bandwidth h, the triangular kernel and limiting observations to those on
    # the same side of the cutoff c. In some sense, this is purely visual, as
    # the IK bandwidth is optimized for x \in [c-h, c+h] and other data are not
    # involved.
    ll <- function(x_i, y, x, h, c) {
        wgts <- (1 - abs(x-x_i) / h) * (abs(x-x_i) <= h) * 
            (sign(x_i-c)==sign(x-c) | x_i==c)
        lm.fitted <- lm(y ~  x , weights = wgts)
        if(sum(wgts>0)>10) { # Require 10 observations around x_i
            return(predict(lm.fitted, newdata=data.frame(x=x_i)))
        } else {
            return(NA)
        }
    }

    df <- data.frame(y, x, c)
    # Add the fitted value to the dataset
    df$y_fitted <- unlist(lapply(x, ll, y, x, h, c))
    
    # Make a plot
    library(ggplot2)
    ggplot(df, aes(x)) + 
        geom_point(aes(y=y, color=x>c)) + 
        geom_line(aes(y=y_fitted, color=x>c))
```

# References
